---
title: Putting it all together
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Putting it all together}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r include=FALSE}
library(ggplot2)
```

```{css echo=FALSE}
p>img:first-child:last-child {
  display: block;
  margin-left: auto;
  margin-right: auto;
  border: none;
}
video {
  display: block;
  padding: 0;
  border: 1px solid #AAA;
  margin: 1em auto;
  width: 100%;
  box-sizing: border-box;
}
```

Armed with the knowledge from the previous sections, we can now build the ggviolin shinytableau extension. Structurally, this will be very close to the Data Summary extension from the previous section, but the ggviolin extension makes use of additional features from Shiny itself.

The preamble is the same, with the addition of `library(ggplot2)`:

```{r eval=FALSE}
# filetype: shinyApp

library(shiny)
library(shinytableau)
library(promises)
library(shinyvalidate)
library(ggplot2)

manifest <- tableau_manifest_from_yaml()
```

The dashboard view's UI differs in that it uses `fillPage`, and notice that the plot's height is explicitly set to 100%. We also define a brush.

```{r eval=FALSE}
ui <- function(req) {
  fillPage(
    plotOutput("plot", height = "100%",
      brush = brushOpts("plot_brush", resetOnNew = TRUE)
    )
  )
}
```

The dashboard view's server function uses a function we haven't discussed before, `tableau_select_marks_by_brush_async`. This does exactly what the name implies: using the selection defined from a `plotOutput` brush, it drives selection on a Tableau worksheet. (This function has several limitations, including only working with ggplot2-based plots and only when the plot and the worksheet in question have their dimensions in common; see the `?tableau_select_marks_by_brush_async` help topic for more details.)

Note also that `tableau_setting()` is called to retrieve `plot_title`, `xvar`, and `yvar` settings; we'll be saving those settings in the config dialog. When writing your own ggplot2 extensions that let the user select what column to map to certain aesthetics, you'll want to follow the pattern shown below, where the character vector (e.g. `xvar`) is passed to `aes()` using `!!as.symbol(...)`.

```{r eval=FALSE}
server <- function(input, output, session) {
  df <- reactive_tableau_data("data_spec")

  observeEvent(input$plot_brush, {
    worksheet <- req(tableau_setting("data_spec")$worksheet)
    tableau_select_marks_by_brush_async(worksheet, input$plot_brush)
  })

  output$plot <- renderPlot({
    plot_title <- tableau_setting("plot_title")
    xvar <- tableau_setting("xvar")
    yvar <- tableau_setting("yvar")

    df() %...>% {
      ggplot(., aes(x = !!as.symbol(xvar), y = !!as.symbol(yvar))) +
        geom_violin(draw_quantiles = c(0.25, 0.5, 0.75)) +
        ggtitle(plot_title)
    }
  })
}
```

The ggviolin config dialog uses the `choose_data` module, just as the Data Summary example did. It also prompts the user for a title, creates a `uiOutput` (that we'll populate with xvar and yvar select inputs), and previews the selected data table using `tableOutput`.

```{r eval=FALSE}
config_ui <- function(req) {
  tagList(
    textInput("title", "Title"),
    choose_data_ui("data", "Choose data"),
    uiOutput("var_selection_ui"),
    tableOutput("preview")
  )
}
```

The config dialog's server function is the most complicated piece. We'll take it a chunk at a time. In the beginning of the function, [shinyvalidate validation rules](https://rstudio.github.io/shinyvalidate/articles/shinyvalidate.html) are added, which ensure that the user provides these values before settings can be saved.

```{r eval=FALSE}
config_server <- function(input, output, session, iv) {
  iv$add_rule("title", sv_required())
  iv$add_rule("xvar", sv_required())
  iv$add_rule("yvar", sv_required())
```

Next, we invoke the `choose_data()` module, passing in the same `"data"` id that we just passed to `choose_data_ui()` a few moments ago.

```{r eval=FALSE}
  data_spec <- choose_data("data", iv = iv)
```

For the ggviolin extension, not only do we want to save this data spec to the settings, we actually need to use it immediately:

First, for preview purposes, we want the first five rows of the selected table, and we use `reactive_tableau_data` for this.

```{r eval=FALSE}
  data <- reactive_tableau_data(data_spec, options = list(maxRows = 5))

  output$preview <- renderTable({
    data()
  })
```

Second, for letting the user choose which of the selected table's columns should map to the x and y dimensions, we need the schema (which contains the fieldnames, among other things), and we use `reactive_tableau_schema` to get it.

Like `reactive_tableau_data`, the return value for `reactive_tableau_schema` is a reactive expression; but unlike `reactive_tableau_data`, this reactive expression is not asynchronous (doesn't return a promise), so there's no need to worry about using `%...>%`.

```{r eval=FALSE}
  schema <- reactive_tableau_schema(data_spec)

  output$var_selection_ui <- renderUI({
    tagList(
      selectInput("xvar", "Dimension", schema()$columns$fieldName),
      selectInput("yvar", "Measure", schema()$columns$fieldName)
    )
  })
```

Finally, we fulfill the final basic requirement for config server functions: we define and return the function that saves the user's input to Tableau settings.

```{r eval=FALSE}
  save_settings <- function() {
    update_tableau_settings_async(
      plot_title = input$title,
      data_spec = data_spec(),
      xvar = input$xvar,
      yvar = input$yvar
    )
  }
  return(save_settings)
}
```

All that's left now is to hand all of these pieces over to `tableau_extension`.

```{r eval=FALSE}
tableau_extension(
  manifest, ui, server, config_ui, config_server,
  options = ext_options(config_width = 600, config_height = 600, port = 2468)
)
```
